Sufficient Conditions for Reachability in Automata Networks with Priorities

### 2015-02-13 parts/def-ph.tex

\begin{definition}[$\Cap : \PHl \times \PHproc \rightarrow \PHl$]
\label{def:statecap}
  Given a state $s \in \PHl$ and a local state $a_i \in \PHproc$, $(s \Cap a_i)$ is the state defined by:
  $\PHget{(s \Cap a_i)}{a} = a_i \wedge \forall b \neq a, \PHget{(s \Cap a_i)}{b} = \PHget{s}{b}$.
  We can also extend this definition to a set $ps$ of local states,
  under the condition that all local states of $ps$ are from different automata,
  by the override of each local state independently:
  $\forall ps \in \PHsublset[\PHl], s \Cap ps = s \underset{a_i \in ps}{\Cap} a_i$.
\end{definition}
% The notation $\toset{ps}$ represents the set of all elements in $ps$
% (which is not a set but a tuple)
% as defined on page~\pageref{notations}.

\begin{definition}[Semantics of an AAN ($\PHPtrans$)]
\label{def:play}
  An action $h = \PHhit{A}{b_j}{b_k} \in \PHa$ is \emph{playable} in $s \in \PHl$
  if and only if $A \subseteq s$ and $\PHget{s}{b} = b_j$.
  In such a case, $(s \PHplay h)$ stands for the state resulting from the play of the action $h$ in $s$, which is defined by: $(s \PHplay h) = s \Cap b_k$.
  Moreover, we denote: $s \PHPtrans (s \PHplay h)$.

@@

We also extend the override operator to contexts (\pref{def:ctxcap}).
\begin{definition}[$\Cap: \Ctx \times \powerset(\PHproc) \rightarrow \Ctx$]
\label{def:ctxcap}
  For any $\ctx\in\Ctx$ and set of local states $ps \in \powerset(\PHproc)$,
  the override of $\ctx$ by $ps$ is noted $\ctx \Cap ps$ and is defined by:
  \[ \forall a \in \PHs, \PHget{(\ctx \Cap ps)}{a} \DEF
  \begin{cases}
    \{ p \in ps \mid \PHsort(p)=a \} & \text{if } \exists p \in ps, \PHsort(p)=a,\\
    \PHget{\ctx}{a} & \text{else.}
  \end{cases}
  \]
\end{definition}

>>>>

\begin{comment}
\begin{definition}[$\Cap : \PHl \times \PHproc \rightarrow \PHl$]
\label{def:statecap}
  Given a state $s \in \PHl$ and a local state $a_i \in \PHproc$, $(s \Cap a_i)$ is the state defined by:
  $\PHget{(s \Cap a_i)}{a} = a_i \wedge \forall b \neq a, \PHget{(s \Cap a_i)}{b} = \PHget{s}{b}$.
  We can also extend this definition to a set $ps$ of local states,
  under the condition that all local states of $ps$ are from different automata,
  by the override of each local state independently:
  $\forall ps \in \PHsublset[\PHl], s \Cap ps = s \underset{a_i \in ps}{\Cap} a_i$.
\end{definition}
\end{comment}
% The notation $\toset{ps}$ represents the set of all elements in $ps$
% (which is not a set but a tuple)
% as defined on page~\pageref{notations}.

\begin{definition}[Semantics of an AAN ($\PHPtrans$)]
\label{def:play}
  An action $h = \PHhit{A}{b_j}{b_k} \in \PHa$ is \emph{playable} in $s \in \PHl$
  if and only if $A \subseteq s$ and $\PHget{s}{b} = b_j$.
  In such a case, $(s \PHplay h)$ stands for the state resulting from the play of the action $h$ in $s$, which is defined by:
    $(s \PHplay h)_b = b_k$
	and
	$\forall a\in\PHs, a\neq b, (s\PHplay h)_a = s_a$.
  Moreover, we denote: $s \PHPtrans (s \PHplay h)$.

@@

\begin{comment}
We also extend the override operator to contexts (\pref{def:ctxcap}).
\begin{definition}[$\Cap: \Ctx \times \powerset(\PHproc) \rightarrow \Ctx$]
\label{def:ctxcap}
  For any $\ctx\in\Ctx$ and set of local states $ps \in \powerset(\PHproc)$,
  the override of $\ctx$ by $ps$ is noted $\ctx \Cap ps$ and is defined by:
  \[ \forall a \in \PHs, \PHget{(\ctx \Cap ps)}{a} \DEF
  \begin{cases}
    \{ p \in ps \mid \PHsort(p)=a \} & \text{if } \exists p \in ps, \PHsort(p)=a,\\
    \PHget{\ctx}{a} & \text{else.}
  \end{cases}
  \]
\end{definition}
\end{comment}

### 2015-02-13 parts/example.tex

At this point, the static analysis results presented in
\pref{ssec:ua} can be used to check if the model is functional,
\ie if gene $a$ can oscillate, thus leading to the formation of stripes.
Starting from context $\ctx^1 = \PHstate{f_1, a_0, c_0}$,
we thus want to check the reachability of $\myp^1 = a_1$;
then, starting from any new state obtained,
we want to check the reachability of $\myp^2 = a_0$,
and finally $\myp^3 = a_1$ once again to ensure that we entered a cycle.
This consists in using \pref{th:approxinf} several times
to check these reachabilities successively.

@@

In conclusion, by using \pref{th:approxinf} on these local causality graphs,
and the traversal of \pref{ssec:concret} to determine
the starting states of the following reachability step,
we showed that it is possible to reach successively
$a_1$, $a_0$ and $a_1$
and thus that the AAN of \pref{fig:metazoan-ph} is functional.
This result can be extended to the model of \pref{fig:metazoan-php}
because they have the same dynamics (given \pref{th:bisimPHP}).

>>>>

At this point, the static analysis results presented in
\pref{ssec:ua} can be used to check if the model is functional,
\ie if gene $a$ can oscillate, thus leading to the formation of stripes.
Starting from context $\ctx^1 = \PHstate{f_1, a_0, c_0}$,
we thus want to check the reachability of $\myp^1 = a_1$;
then, starting from any new state obtained,
we want to check the reachability of $\myp^2 = a_0$,
and finally $\myp^3 = a_1$ once again to ensure that we entered a cycle.
This consists in using \pref{th:approxinf} several times
to check these reachabilities successively,
by splitting it following \pref{ssec:ordered-ua}.

@@

In conclusion, by following \pref{ssec:ordered-ua},
we showed that it is possible to reach successively
$a_1$, $a_0$ and $a_1$
and thus that the AAN of \pref{fig:metazoan-ph} is functional.
This result can be extended to the model of \pref{fig:metazoan-php}
because they have the same dynamics (given \pref{th:bisimPHP}).

### 2015-02-13 parts/models-ph-php.tex
revised section 2
remove section 4.1 (ADN to AAN)
fix section mistake
remove \mod commands

\subsection{Asynchronous Automata Networks with classes of priorities}
\label{ssec:flattening}

@@

\begin{definition}[Dynamics of an AAN$k$ ($\PHPtrans$)]
\label{def:playp}
  An action $h = \PHhit{A}{b_j}{b_k} \in \PHa^{(n)}$ of priority $n$
  is \emph{playable} in $s \in \PHl$
  if and only if $A \subseteq s$, $\PHget{s}{b} = b_j$ and
  $\forall m < n, \forall g \in \PHa^{(m)},
    \neg (\PHhitter(g) \subseteq s \wedge \PHtarget(g) \in s)$.
  In such a case, $(s \PHplay h)$ stands for the state resulting from the play
  of the action $h$ in $s$, which is defined by: $(s \PHplay h) = s \Cap b_k$.
  Moreover, we denote: $s \PHPtrans (s \PHplay h)$.
\end{definition}

@@

\begin{definition}[Playability property operator ($\Fopsymbol : \PHh \rightarrow \F$)]\label{def:fop}
  For all $h = \PHfrappe{A}{b_j}{b_m} \in \ov{\PHh}$, we define:
  \[
    \Fop{h} \equiv
    b_j \wedge
    \left(\bigwedge_{a_i \in A} a_i\right)
    %\hitter{h}
    \wedge
      \left( \bigwedge_{\substack{g \in \ov{\PHh}^{(n)}\\\modMF{1 <} n < \prio(h)}}
      \neg \left( \target{g} \wedge \left(\bigwedge_{c_l \in \hitter{g}} c_l\right)\right) \right)
%   \moda{\bigwedge_{\substack{b_z \in \PH_b\\b_z \neq b_x}} \neg b_z}
  \]
\end{definition}
%
By construction of this operator and given the dynamics of a AAN$k$,
an action $h$ is playable in a state $s \in \ov{\PHl}$ if and only if: $\Fsem{\Fop{h}}{s}$.

>>>>

\section{Asynchronous Automata Networks with classes of priorities}
\label{sec:flattening}

@@

\begin{definition}[Semantics of an AAN$k$ ($\PHPtrans$)]
\label{def:playp}
  An action $h = \PHhit{A}{b_j}{b_k} \in \PHa^{(n)}$ of priority $n$
  is \emph{playable} in $s \in \PHl$
  if and only if $A \subseteq s$, $\PHget{s}{b} = b_j$ and
  $\forall m < n, \forall g \in \PHa^{(m)},
    \neg (\PHhitter(g) \subseteq s \wedge \PHtarget(g) \in s)$.
  In such a case, $(s \PHplay h)$ stands for the state resulting from the play
  of the action $h$ in $s$, which is defined by: 
    $(s \PHplay h)_b = b_k$
	and
	$\forall a\in\PHs, a\neq b, (s\PHplay h)_a = s_a$.
  Moreover, we denote: $s \PHPtrans (s \PHplay h)$.
\end{definition}

@@

\begin{definition}[Playability property operator ($\Fopsymbol : \PHh \rightarrow \F$)]\label{def:fop}
  For all $h = \PHfrappe{A}{b_j}{b_m} \in \ov{\PHh}$, we define:
  \[
    \Fop{h} \equiv
    b_j \wedge
    \left(\bigwedge_{a_i \in A} a_i\right)
    %\hitter{h}
    \wedge
      \left( \bigwedge_{\substack{g \in \ov{\PHh}^{(n)}\\1 < n < \prio(h)}}
      \neg \left( \target{g} \wedge \left(\bigwedge_{c_l \in \hitter{g}} c_l\right)\right) \right)
  \]
\end{definition}
%
By construction of this operator and given the dynamics of a AAN$k$,
an action $h$ is playable in a state $s \in \ov{\PHl}$ if and only if: $\Fsem{\Fop{h}}{s}$.

### 2015-02-13 parts/sa.tex

\todo{name + prototype of the procedure}

\todo{example}

>>>>

The execution of such a traversal outputs a scenario.
We denote $\Delta(\cwB)$ the set of all scenarios extracted from the
local causality graph $\cwB$.
Indeed, several scenarios may exist for a same graph,
generated by the successors that can be arbitrarily chosen
in this algorithm.
An example of such a traversal is given on a small example
in \pref{ssec:ex-metazoan}.
